Almost orthogonal vectors

### 2014-12-08 project3/paper/paper.tex
Added annealing description to paper.

\proof
Let $\delta= \max_{i\neq j}\{v_i\cdot v_j\}$ for a given set of unit vectors $v_1, v_2, \ldots, v_{n+1}$, and let $\delta_{min}$ be the minimum possible $\delta$ obtained across all unit vectors $v_1, v_2, \ldots, v_{n+1}$. We first show that $\delta_{min} \le-\frac{1}{n}$ by showing that the origin-centered simplex configuration of the vectors $v_1,\cdots v_{n+1}$ produces a value $\delta=-\frac{1}{n}$, and then will later show that $\delta_{min} \geq -\frac{1}{n}$ as well.

@@

Since $\delta_{min} \le -\frac{1}{n}$ and $\delta_{min} \ge -\frac{1}{n}$, we can conclude that $\delta_{min}=-\frac{1}{n}$. From this, we can see that a $\epsilon = \frac{1}{n^2}$ is achievable for the general case of when $N=n+1$ -- this leads us to conclude that the minimum possible $\epsilon$ value for $N=n+1$ is at most $\frac{1}{n^2}$.
\section{$N>n+1$}

We conjecture that in cases where $N>n+1$ we can never have an arrangement of the vectors where all values of $|v_i\cdot v_j|^2$ are the same. Proving this remains an object of future study.

\section{Applications to sphere packing}
We should be able to prove something about the radius of spheres you can pack in the case of $N=n+1$, taking as intuition square and hexagonal tilings of the plane and cubic and tetrahedral tilings of 3-space.

>>>>

\begin{proof}
Let $\delta= \max_{i\neq j}\{v_i\cdot v_j\}$ for a given set of unit vectors $v_1, v_2, \ldots, v_{n+1}$, and let $\delta_{min}$ be the minimum possible $\delta$ obtained across all unit vectors $v_1, v_2, \ldots, v_{n+1}$. We first show that $\delta_{min} \le-\frac{1}{n}$ by showing that the origin-centered simplex configuration of the vectors $v_1,\cdots v_{n+1}$ produces a value $\delta=-\frac{1}{n}$, and then will later show that $\delta_{min} \geq -\frac{1}{n}$ as well.

@@

Since $\delta_{min} \le -\frac{1}{n}$ and $\delta_{min} \ge -\frac{1}{n}$, we can conclude that $\delta_{min}=-\frac{1}{n}$. From this, we can see that a $\epsilon = \frac{1}{n^2}$ is achievable for the general case of when $N=n+1$ -- this leads us to conclude that the minimum possible $\epsilon$ value for $N=n+1$ is at most $\frac{1}{n^2}$.
\end{proof}

Note that proving that the minimum possible epsilon for the case of $N = n+1$ is equal to $1/n^2$ is much harder, and we leave this for future work.

\section{Simulated annealing and associated conjectures}
Since visualizing in spaces of dimensionality greater than $3$ is extremely hard, we used an annealing approach to compute near-optimal epsilon values for different $(n, N)$ pairs. We describe this method in greater detail below.

Our general intuition tells us that perturbing a set of nearly optimal vectors slightly could give us a new set of vectors that produce even a smaller epsilon value. Given this, we start off with a set of random vectors, and try to move these vectors in order to produce smaller epsilon values.

At every iteration, we maintain the set of vectors (called cur\_vectors) that have produced the optimal epsilon value so far. Then we compute a set new\_vectors defined as follows,
$$\text{new\_vector} = \text{old\_vector} + (1/i \cdot \text{random\_vector})$$

We update cur\_vectors to the set new\_vectors if the epsilon value produced by new\_vectors is less than the epsilon value produced by cur\_vectors. $i$ is incremented whenever cur\_vectors is updated -- having the $1/i$ factor in the above expression ensures that as we move closer to an optimal value, we perturb the set of vectors by a smaller amount.

Even though this approach seems simple, it actually produces remarkably precise values -- this can be verified by cross-checking with results we've already proven / conjectured in this paper.

To make this more concrete, consider Table $1$.
\begin{table}
   \centering
    \begin{tabular}{ | c | c | c | c |}
    \hline
      & $n=2$ & $n=3$ & $n=4$ \\ \hline
    $N=2$ & $0.0$ & $-$ & $-$ \\ \hline
    $N=3$ & $0.25$ & $0.0$ & $-$ \\ \hline
    $N=4$ & $0.5$ & $0.11$ & $0.0$ \\ \hline
    $N=5$ & $0.66$ & $0.21$ & $0.07$ \\ \hline
    $N=6$ & $0.75$ & $0.22$ & $0.14$ \\ \hline
    $N=7$ & $0.82$ & $0.36$ & $0.18$ \\ \hline
    $N=8$ &  $0.86$ & $0.44$ & $0.24$ \\ \hline
    $N=9$ &  $0.89$ & $0.49$ & $0.26$ \\ \hline
    $N=10$ & $0.91$ & $0.51$ & $0.32$ \\ \hline
    $N=11$ & $0.93$  & $0.55$ & $0.35$ \\ \hline
    $N=12$ & $0.94$ & $0.64$ & $0.41$ \\
    \hline
    \end{tabular}
    \caption {Minimum values of $\epsilon$ for different values of $(n, N)$, as computed by our annealing method}
\end{table}

The accuracy of our method can be noted by looking at the computed values for $n=2$ -- the computed values for different values of $N$ are very close to the expected value of $\cos^2 \pi/N$.
\section{Applications to sphere packing}
We should be able to prove something about the radius of spheres you can pack in the case of $N=n+1$, taking as intuition square and hexagonal tilings of the plane and cubic and tetrahedral tilings of 3-space.
